\LARGE \bf Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation of RGB-D Images

### 2013-09-14 paper.tex
squeezing and polishing
git-svn-id: https://svn.ais.uni-bonn.de/svn/staff/publications/icra_2014_mueller_nyu@4379 b2511e17-9ac0-488b-bc21-1d6d82e5baa2

\section{RELATED WORK}
The task of dense semantic annotation of 3D maps has seen an increased interest
in recent years.  Early work includes \citet{nuchter2008towards}, who combine
6D SLAM, surface annotation and object recognition, to build semantically
annotated maps. They demonstrate their approach on a mobile robot in an indoor
environment.
%
More recently \citet{sengupta2012automatic} introduced a dataset of
semantically annotated street-scenes in a closed track, captured as pairs of
stereo images. They approach the task by jointly reasoning about 3D layout and
semantic of the scenes and produce a dense labeling on image level.  In
\citet{senguptaurban}, they extend their approach to produce a volumetric
reconstruction of the scene, together with a dense semantic labeling of the
volumetric representation. Their approach to image segmentation is build on
the hierarchical CRF apporach of
\citet{ladicky2009associative}, which is similar in spirit to our approach,
but uses Potts potentials together with cross-validation to set potentials.

Recent approaches for indoor semantic annotation of maps mostly focused on
RGB-D images, which are now easy to obtain using structured light sensors.
\citet{stuckler2012semantic} used a Random Forest model to obtain a dense
semantic labeling of images, integrating predictions over multiple views in 3D.
They evaluated their approach on a limited set of table-top and indoors scenes.
%
\citep{silberman2011indoor} introduced the NYU Depth Dataset V1 indoor dataset,
which consisted of a large variety of densely annotated indoor scences, captured as RGB-D images.
Their work also introduced a baseline method for semantic segmentation of RGB-D
image, which used an approach based on superpixels, interest point descriptors
and a CRF.
While pairwise potentials for the CRF were carefully
designed for the dataset, potentials were either directly set by hand or estimated using empirical
frequencies. This is in contrast to our work, which applies structured prediction techniques
to learn potentials automatically to optimize predictive performance.
%
\citet{ren2012rgb} evaluated the design of features for semantic labeling of RGB-D data, and use
a hierarchical segmentation to provide context. While they also define a CRF on superpixels,
their model is again not learned, but a weighted Potts model, using only a
probability of boundary map, and not taking spatial layout into account at all.
%
\citet{SilbermanECCV12} extended the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
using in this work.  Their focus is on inferring support relations in indoor
scenes, such as objects resting on tables or shelfs, which in turn rest on the floor.
Their approach is based on robust estimation of 3D plane hypotheses, which are then jointly
optimized with support relations and structure classes.
\citet{SilbermanECCV12} use a complex pipeline, employing significant domain knowledge.
In this work, on the other hand, we try to learn all relevant domain specific features
directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
with respect to structure class segmentation.

@@

While many of the works mentioned in this chapter make use of a CRF
approach, we are not aware of any prior work on semantic annotation of 3D maps
that fully learns their potentials. This allows our approach to infer
the scene structure from the data, without the need to specify relevant relations
by hand.
%The task of dense semantic labeling in natural images has been widely studied
%in computer vision.  Our approach of using pixel-based low-level cues is
%similar to the work of \citet{shotton2006textonboost}, who introduced
%TextonBoost for semantic segmentation, and used a hand-coded CRF energy to
%model spatial interactions. More recent work usually employ learned conditional
%random fields, learning object interactions from the
%data to directly maximize a loss function~\citep{lucchi2013learning, krahenbuhl2012efficient},
%which is the approach we are following.

@@

\subsection{Unary Image Features}\seclabel{unaries}
Our method builds on the probability output of a random forest, trained for pixel-wise
classification of the structure classes.
We use the GPU implementation provided by \citet{stueckler2013}\footnote{\url{https://github.com/deeplearningais/curfil}}.
The input for training are the full RGB-D images, transformed to Lab
color space. Each tree in the forest uses training pixels only from a subset of
training images. For each training image, an equal number of pixels for each
occurring class is sampled. Split features are given by difference of regions
on color or depth channels. Region size and offsets are normalized using the
depth at the target pixel.
We accumulate the probabilistic output for all pixels within a superpixel, and
use the resulting distribution as a feature for the unary node potentials in our CRF model.
We augment these prediction with another feature, based on the height of a superpixel in 3D.
This is a very informative feature, in particular to determine the floor.
To compute the height of a (super) pixel, we first find the ``up'' direction.
We use a very simple approach that we found effective: we cluster normal
directions of all pixels into 10 clusters using $k$-means, and use the one that
is most parallel to the $Y$ direction, which roughly corresponds to height in
the dataset.
We then project the 3D point cloud given by the depth along this direction, and
normalize the result between $0$ and $1$.
This procedure works robustly given there is some horizontal surface in the image,
such as the ground or a table. A few scenes contain only walls and furniture,
and the approach fails for these. \Figref{height} illustrates a typical case and
one of the much rarer failure cases.
While we could use a more elaborate scheme,
such as the one from \citet{SilbermanECCV12}, we suspect that the feature is of
little use in scenes without horizontal surfaces.
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{images/normal_feature}
    \end{center}
    \caption{%
        Visualization of one of the pairwise features, the similarity between superpixel normals.
        The image shows the zoom-in of a bedroom scene, together with the superpixel over-segmentation.
        Lines connect adjacent superpixels, and line-strength gives the magnitude of the orientation similarity.
    \figlabel{normal_feature}}
\end{figure}

@@

\emph{Color Contrast}
    We employ a non-linear color contrast, as is common in the computer
    vision literature, between the superpixel mean colors $c_i$ and $c_j$:
    \[\exp\left(-\gamma \lVert c_i - c_j \rVert^2\right).\]

@@

\bottomrule
\end{tabularx}
\quad\\The best value in each column is printed in bold$^\dagger$. The upper part of
the table shows contributions by different parts of our pipeline. RF stands for random forest prediction, RF + SP for aggregated
random forests prediction within superpixels, RF + SP + SVM for an SVM trained on the unary potentials, and RF + SP + CRF is
our proposed pipeline. We optimized our approach for class average
accuracy.\\
$^\dagger$ \footnotesize Note that the work of \citet{stueckler2013} is not directly
comparable, as they integrated information over multiple frames, and did not
measure accuracy for pixels without valid depth measurement.
\end{table*}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=.45\linewidth]{images/vertical_alignment}\vspace{1mm}
    \includegraphics[width=.45\linewidth]{images/depth_difference}
    \end{center}
    \caption{%
        Visualization of some of the learned potentials. The right potential is
        applied to relative depth between superpixels, the second on the
        feature encoding whether one superpixel is above the other in the
        image. See \secref{results} for details.
    \figlabel{weights}}
\end{figure}

@@

\section{SUMMARY AND DISCUSSION}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
of random forest, and model spatial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function that is fully learned using 
 the SSVM approach. By explicitly modeling 3D
 relations in a fully learned framework, we improve the state-of-the art on the
 NYU V2 dataset for semantic annotation of structure classes.

>>>>

\section{RELATED WORK}
The task of dense semantic annotation of 3D maps has seen an increased interest
in recent years.  Early work includes \citet{nuchter2008towards}, who combine
6D SLAM, surface annotation, and object recognition to build semantically
annotated maps. They demonstrate their approach on a mobile robot in an indoor
environment.
%
More recently \citet{sengupta2012automatic} introduced a dataset of
semantically annotated street-scenes in a closed track, captured as pairs of
stereo images. They approach the task by jointly reasoning about 3D layout and
semantics of the scenes and produce a dense labeling on image level.  In
\citet{senguptaurban}, they extend their approach to produce a volumetric
reconstruction of the scene, together with a dense semantic labeling of the
volumetric representation. Their approach to image segmentation is build on
the hierarchical CRF approach of
\citet{ladicky2009associative}, which is similar in spirit to our approach,
but uses Potts potentials together with cross-validation to set potentials.

Recent approaches for indoor semantic annotation of maps mostly focused on
RGB-D images, which are now easy to obtain using structured light sensors.
\citet{stuckler2012semantic} used a Random Forest model to obtain a dense
semantic labeling of images and integrated predictions over multiple views in 3D.
They evaluated their approach on table-top and simple indoors scenes.
%
\citet{silberman2011indoor} introduced the NYU Depth Dataset V1 indoor dataset,
which consisted of a large variety of densely annotated indoor scenes, captured as RGB-D images.
Their work also introduced a baseline method for semantic segmentation of RGB-D
image, which is based on a CRF over superpixels, with unary potentials given by interest point descriptors.
While pairwise potentials for the CRF were carefully
designed for the dataset, potentials were either directly set by hand or estimated using empirical
frequencies. This is in contrast to our work, which applies structured prediction techniques
to learn potentials automatically to optimize predictive performance.
%
\citet{ren2012rgb} evaluated the design of features for semantic labeling of
RGB-D data, and use a hierarchical segmentation to provide context. While they
also define a CRF on superpixels, their model is again not learned, but a
weighted Potts model, using only a probability of boundary map, and not taking
spatial layout into account at all.
%
\citet{SilbermanECCV12} extended the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
using in this work.  Their focus is on inferring support relations in indoor
scenes, such as objects resting on tables or shelfs, which in turn rest on the floor.
Their approach is based on robust estimation of 3D plane hypotheses, which are then jointly
optimized with support relations and structure classes.
\citet{SilbermanECCV12} use a complex pipeline, employing significant domain knowledge.
In this work, on the other hand, we try to learn all relevant domain specific features
directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
with respect to structure class segmentation.

@@

While many of the works mentioned in this chapter make use of a CRF
approach, we are not aware of any prior work on semantic annotation of 3D maps
that fully learns their potentials.
%The task of dense semantic labeling in natural images has been widely studied
%in computer vision.  Our approach of using pixel-based low-level cues is
%similar to the work of \citet{shotton2006textonboost}, who introduced
%TextonBoost for semantic segmentation, and used a hand-coded CRF energy to
%model spatial interactions. More recent work usually employ learned conditional
%random fields, learning object interactions from the
%data to directly maximize a loss function~\citep{lucchi2013learning, krahenbuhl2012efficient},
%which is the approach we are following.

@@

\subsection{Unary Image Features}\seclabel{unaries}
Our method builds on the probability output of a random forest, trained for pixel-wise
classification of the structure classes.
We use the GPU implementation provided by \citet{stueckler2013}\footnote{\url{https://github.com/deeplearningais/curfil}}.
The input for training are the full RGB-D images, transformed to Lab
color space. Each tree in the forest uses training pixels only from a subset of
training images. For each training image, an equal number of pixels for each
occurring class is sampled. Split features are given by difference of regions
on color or depth channels. Region size and offsets are normalized using the
depth at the target pixel.
We accumulate the probabilistic output for all pixels within a superpixel, and
use the resulting distribution as a feature for the unary node potentials in our CRF model.
We augment these prediction with another feature, based on the height of a superpixel in 3D.
This is a very informative feature, in particular to determine the floor.
To compute the height of a (super) pixel, we first find the ``up'' direction.
We use a very simple approach that we found effective: we cluster normal
directions of all pixels into 10 clusters using $k$-means, and use the one that
is most parallel to the $Y$ direction, which roughly corresponds to height in
the dataset.
We then project the 3D point cloud given by the depth along this direction, and
normalize the result between $0$ and $1$.
This procedure works robustly given there is some horizontal surface in the image,
such as the ground or a table. A few scenes contain only walls and furniture,
and the approach fails for these. \Figref{height} illustrates a typical case and
one of the much rarer failure cases.
While we could use a more elaborate scheme,
such as the one from \citet{SilbermanECCV12}, we suspect that the feature is of
little use in scenes without horizontal surfaces.
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{images/normal_feature}
    \end{center}
    \vspace{-3mm}
    \caption{%
        Visualization of one of the pairwise features, the similarity between superpixel normals.
        The image shows the zoom-in of a bedroom scene, together with the superpixel over-segmentation.
        Lines connect adjacent superpixels, and line-strength gives the magnitude of the orientation similarity.
    \figlabel{normal_feature}}
    \vspace{-9mm}
\end{figure}

@@

\emph{Color Contrast}
    We employ a non-linear color contrast, as is common in the computer
    vision literature, between the superpixel mean colors $c_i$ and $c_j$:
    $\exp\left(-\gamma \lVert c_i - c_j \rVert^2\right)$.

@@

\bottomrule
\end{tabularx}
\quad\\The best value in each column is printed in bold$^\dagger$. The upper part of
the table shows contributions by different parts of our pipeline. RF stands for random forest prediction, RF + SP for aggregated
random forests prediction within superpixels, RF + SP + SVM for an SVM trained on the unary potentials, and RF + SP + CRF is
our proposed pipeline. We optimized our approach for class average
accuracy.\\
$^\dagger$ \footnotesize Note that the work of \citet{stueckler2013} is not directly
comparable, as they integrated information over multiple frames, and did not
measure accuracy for pixels without valid depth measurement.
\vspace{-5mm}
\end{table*}

@@

\begin{figure}
    \begin{center}
        \includegraphics[width=.45\linewidth]{images/vertical_alignment}
    \includegraphics[width=.45\linewidth]{images/depth_difference}
    \end{center}
    \vspace{-3mm}
    \caption{%
        Visualization of some of the learned potentials. The right potential is
        applied to relative depth between superpixels, the second on the
        feature encoding whether one superpixel is above the other in the
        image. See \secref{results} for details.
    \figlabel{weights}}
    \vspace{-3mm}
\end{figure}

@@

\section{SUMMARY AND DISCUSSION}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
of random forest, and model spatial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function
that is learned using the SSVM approach. By explicitly modeling 3D relations in
a fully learned framework, we improve the state-of-the art on the NYU V2
dataset for semantic annotation of structure classes.
